---
title: "Homework 4"
author: "Rylee Pupa"
format: docx
editor: visual
execute: 
  warning: false
  message: false
---

```{r}
library(here)
library(janitor)
library(tidyverse)
library(naniar)
library(visdat)
library(tidyverse)
library(here)
library(lterdatasampler)
library(performance)
library(broom)
library(flextable)
library(ggeffects)
library(car)
```

1.  Write your null and alternative hypotheses in mathematical and biological terms.
    -   null: fish length does not predict fish weight for trout perch across all sample years
        -   math: In statistical terms, r = 0
    -   alternative: fish length does predict fish weight for trout perch across all sample years
        -   math: In statistical terms, r is not equal to zero

```{r}
#reading in files using here function, clean names to make lowercase and more readable
fish <- read_csv(here("data", "ntl6_v12.csv")) %>% 
        clean_names()

#filter fish to only include trout perch
trout_perch <- fish %>% 
        filter(spname == "TROUTPERCH")
```

```{r}
#2.visualize missing data
gg_miss_var(trout_perch) + labs(x = "Missing data (#)",
       y = "Variables", caption = "The missing weight data is relevant to the hypothesis")
#Extended Caption: There is missing data for sex, flag, fish part, sample, type, and weight. There are 200 missing data values for weight, which is the only missing data relevant to our hypothesis. 
```

```{r exploratory visuals}
#plotting points to explore if check if there may be a linear relationship in the data
ggplot(data = trout_perch, aes(x = length, y = weight)) +
    geom_point()
```

```{r}
#3 running our linear tests (can check assumptions from this)
#creating a linear model of the length and weight
troutperch_model <- lm(weight ~ length, data = trout_perch)

#make pane show grid of diagnostic plots from linear model 
par(mfrow = c(2,2))
plot(troutperch_model) 
  
#to turn off two by two grid 
#dev.off()

#remove outliers
```

#5) For each diagnostic plot, describe in 1-2 sentences what it is showing you, and what youdecide after looking at the plot. Check assumptions:

1.  The Residuals vs. fitted plot shows us if the errors or residuals in the relationship are homoscedastic, in other words whether or not there is a constant variance. This is visible because the errors plotted are mostly constant throughout the range of x-axis values, most except outliers are evenly and randomly distributed across the horizontal line
2.  The Normal QQ plot shows us if the errors are normally distributed, with the straight line through the plot I would assume yes it is normally distibuted.
3.  The scale location plot also checks for homoscedasticy of variance but instead uses square root of standardized residuals. The variance of residuals is evenly and randomly distributed so I would also say that the residuals are homockedastic.
4.  The residuals versus leverage plot (also called Cook's model) helps us check if outliers are influencing our model estimates by measuring the influence of a single observation on the model. Since there are several points labeled in the graph, I would say yes outliers are influencing.

Other assumptions not shown in grid:

Assumption 1: There is a linear relationship between variables: Yes! From our exploratory data visualization we can see a mostly linear relationship despite outliers.

Assumption 2: The errors are independent (no correlation): Yes we would assume so because the data collected was independent, so we can assume that errors in one measurement wouldn't influence errors in other observations. 

```{r}
#storing our summary model as an object 
model_summary <- summary(troutperch_model)

#storing ANOVA table as an object
model_anova <- anova(troutperch_model)

#6) displaying results from summary using the model object
model_summary

model_anova
```

orsummarizing model in a table

```{r}
#7 creating a table that summarizes the anova table 

model_anova_table <- tidy(model_anova) %>% 
    #rounding the statistics, sum of mean squares and sum of squares, to two digits
    mutate(across(sumsq:meansq, ~ round(.x, digits = 2))) %>% 
    #round the F-statistic value to have 2 digits
    mutate(statistic = round(statistic, digits = 2)) %>% 
    ##replace the p value with < 0.001 rather than a super small number
    mutate(p.value = case_when(p.value < 0.001 ~ "< 0.001")) %>% 
    #rename the length to be more proper
    mutate(term = case_when(term == "length" ~ "Length", TRUE ~ term)) %>% 
    #create a flex table object from the data frame
    flextable() %>% 
    # change the header labels to be meaningful??
    set_header_labels(df = "Degrees of Freedom", 
                    sumsq = "Sum of Squares",
                    meansq = "Mean Squares",
                    statistic = "F-statistic",
                    p.value = "P-value")

model_anova_table
```

7.  In 1-2 sentences, describe how the ANOVA table relates to the information you get from the **summary()** object

8.  In 2-3 sentences, summarize your results in prose with in-text references to test results. Include all relevant information.

```{r}
#10) creating visualization with model predictions and confidence intervals 

#find model predictions for the linear model using ggpredict 
troutperch_predict <- ggpredict(troutperch_model, terms = "length")
troutperch_predict

#plotting predictions from the ggpredict, with length on x and weight on y
predict_plot <- ggplot(data = troutperch_model, 
                       aes(x = length, y = weight)) +
  #plotting the troutperch data lone
  geom_point() +
  #plotting predictions of linear model, changing aesthetics of plot line
  geom_line(data = troutperch_predict, 
              aes(x = x, y = predicted), 
              color = "red", linewidth = 0.8) +
  #plotting the confidence interval from the gg predict function (95%)
  geom_ribbon(data = troutperch_predict, 
              aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +

  #theme and meaningful labels, expand on caption
  theme_classic() +
  labs(x = "Length",
       y = "Weight", caption = "Model predictions for if fish length predicts fish weight")

predict_plot
```
